Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression

### 2018-07-09 revision_1/paper.tex
gathering all revision in order to work on paper again

\documentclass[11pt]{article}
%\documentclass[ba]{imsart}

@@

\newcommand{\bx}{\mbox{\boldmath $x$}}
\newcommand{\bX}{\mbox{\boldmath $X$}}
\newcommand{\bu}{\mbox{\boldmath $u$}}
\newcommand{\bc}{\mbox{\boldmath $c$}}
\newcommand{\bs}{\mbox{\boldmath $s$}}
\newcommand{\by}{\mbox{\boldmath $y$}}
\newcommand{\bz}{\mbox{\boldmath $z$}}
\newcommand{\bv}{\mbox{\boldmath $v$}}
\newcommand{\bb}{\mbox{\boldmath $b$}}
\newcommand{\bzero}{\mbox{\boldmath $0$}}
\newcommand{\thstarj}{\mbox{$\theta^\star_j$}}
\newcommand{\bths}{\mbox{$\btheta^\star$}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

@@

\title{Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression}
\author{John R. Lewis, Steven N. MacEachern and  Yoonkyung Lee \\
{\small \it Department of Statistics, The Ohio State University, Columbus, Ohio 43210}\\
{\small lewis.865@osu.edu, snm@stat.osu.edu and yklee@stat.osu.edu}
\thanks{This research has been supported by Nationwide Insurance Company and by the NSF under grant numbers DMS-1007682 and DMS-1209194.  The views in this paper are not necessarily those of Nationwide Insurance or the NSF.}} 
\begin{document}
\date{}
\maketitle

\begin{abstract}
The \textit{restricted likelihood} is introduced as a method to address concerns about the miss-specification of the likelihood in Bayesian models. To apply, the data is summarized through a set of insufficient statistics,
targeting inferential quantities of interest, and the prior is updated with only the summary statistics. By a careful choice of the summary, we retain the main benefits of Bayesian methods while reducing the
sensitivity of the analysis to features of the data not captured by
the conditioning statistics. Examples specifically address reducing sensitivity to outliers,
where classical robust estimators (e.g., M-estimators) are natural choices
for conditioning statistics. With these choices, the method can be thought of as a blend of classical robust estimation and Bayesian methods.

A Markov chain Monte Carlo (MCMC) algorithm, applicable for the linear model and a wide range of choices for
summary statistics, is developed to overcome implementation challenges. The method is demonstrated with several examples including simulations and messy insurance agency data with many outliers. Success is manifested in better predictive
performance for data points of interest as compared to competing methods.

@@

\section{Introduction}
Bayesian methods are useful to a wide range of scientific problems, with their value
having been demonstrated both empirically and theoretically. However, these methods run into difficulties
for two major and prevalent classes of problems: handling data sets
with outliers and dealing with model misspecification. From Bayesian perspective the model refers to the prior, the likelihood, as well as the loss function when formal inference is required. While optimality of Bayesian methods is unquestioned if one accepts the validity of all three of these elements, healthy skepticism encourages us to question each.  Concern about the prior distribution has
been addressed through objective Bayesian methods \citep{berger2006} prior elicitation methods \citep{garthwaite2005, ohagan2006}. Concern about the loss function is reflected in the extensive literature on Bayesian hypothesis tests \citep{kass1995}.  Despite the common use of flat-tailed distributions \citep{berger1994} or nonparametric techniques \red{References?}, the sampling density has been given less attention from a specifically Bayesian view. The work on predictive diagnostics \citep{box1980} departs from classical
traditions.

The focus of this work is the development of techniques to handle imperfections in the sampling density.  These imperfections often show
themselves through the presence of outliers--cases not reflecting the phenomenon under
study. There are three main solutions to Bayesian outlier-handling.  The first is to replace the basic sampling
density with a mixture model which includes one component for the ``good'' data and a second 
component for the ``bad'' data.  The second approach replaces the
basic sampling density with a thick-tailed density in an attempt to discount outliers.  The third approach fits a flexible (typically nonparametric) model to the data, producing a Bayesian version of a density estimate for both good and bad data.  In recent 
development, inference is made through the use of robust inference functions \citep[][in press]{lee2013}.

The traditional strategies for handling outliers all have their drawbacks.  While we view the sampling 
density for the good data as stable, the outlier-generating processes 
may be transitory in nature, constantly shifting as the source of bad data changes.  This prevents
us from appealing to large-sample arguments to claim that, with enough data, we can nail down a
model for both good and bad data combined.  Instead of attempting to model both good and bad
data, we propose a new strategy. In a nutshell, we begin with a complete model   
as if all of the data are good. But rather than driving the move from prior distribution to posterior
distribution by the entire likelihood, we use only the likelihood driven by a
few summary statistics that target inferential quantities
of interest.  We call this the \textit{restricted likelihood}. The update is a formal update from prior distribution to posterior distribution, based on the sampling density of the summary statistics.

While the focus of this work is handling outliers, model miss-specification is more general. By focusing on robustly estimating quantities of interest and conditioning on these estimates, the restricted likelihood offers promiss in handling more general miss-specification issues in model building. The remainder of the paper.....develops Bayesian restricted likelihood (Section~\ref{restrictedlikelihood}), shows
how it can be applied to a Bayesian linear model (Section~\ref{BayesLinMod}), illustrates its use
on an insurance agencies data set with a novel twist on model evaluation (Section~\ref{Applications}), 
and wraps up with a discussion (Section~\ref{Conclusions}).  
A major contribution of this work is the computational strategy whose legitimacy is established 
in Section~\ref{BayesLinMod}.  The technical proofs are in the appendix.

\section{The Restricted Likelihood}
\label{restrictedlikelihood}

This sections describes the restricted likelihood using a pair of simple examples and the relates the method to the current literature. For the examples assume $\by=(y_1,\ldots,y_n)$ is a random sample
of size $n$ from a continuous distribution indexed by a parameter
vector $\bth$, with pdf $f(y| \bth)$.  The full-data, likelihood is $L(\bth | \by) = \prod_{i=1}^n f(y_i | \bth)$.

For the first example, consider that a known subset of the data is not informative about $\bth$.  This case mimics the setting where outlying observations  are identified and discarded before formal analysis. 
Labeling the good cases $1$ through $n-k$ and the bad cases $n-k+1$ through $n$.  
The relevant likelihood to be used to move from prior distribution to posterior distribution is 
$L(\bth | y_1, \ldots, y_{n-k}) = \prod_{i=1}^{n-k} f(y_i | \bth)$.  
Equivalently,  the full 
likelihood is rewritten as 
\begin{eqnarray}
\label{OutlyingCases}
L(\bth | \by)  
= \left( \prod_{i=1}^{n-k} f(y_i | \bth) \right) \left( \prod_{i=n-k+1}^{n} f(y_i | \bth) \right) .  
\end{eqnarray}
The second piece, which is independent of $\theta$,  is dropped to provide better inference on $\bth$.

A second example involves deliberate censoring of small and large observations.  This is
sometimes done as a precursor to the analysis of reaction time
experiments  \citep[e.g.,][]{ratcliff1993} where very short and long reaction times are usually explained as anticipation and inattention, respectively, and do not reflect the process being studied.
With lower and upper censoring times at $t_1$ and $t_2$, the post-censoring sampling distribution
is of mixed form, with masses $F(t_1|\bth)$ at $t_1$ and $1-F(t_2|\bth)$ at $t_2$,
and density $f(y | \bth)$ for $y \in (t_1, t_2)$.  Adjusting $y_i$,
with $c(y_i)$ where $c(y_i)= t_1$ if $y_i \leq t_1$, $c(y_i)=t_2$ 
if $y_i \geq t_2$, and $c(y_i)=y_i$ otherwise.  
The adjusted update is performed with $L(\bth |c(\by))$.  
With slightly non-standard notation, we let $g(t_1|\bth) = F(t_1|\bth)$,
$g(t_2 | \bth) = 1 - F(t_2|\bth)$, and $g(y|\bth)=f(y|\bth)$ for
$y \in (t_1, t_2)$. The full likelihood as the product of two pieces, 
\begin{eqnarray}
\label{Censoring}
L(\bth | \by) =  \left( \prod_{i=1}^n g(c(y_i)  | \bth) \right) \left( \prod_{i=1}^n f(y_i|\bth,c(y_i) ) \right) ,  
\end{eqnarray}
and only the first for the formal update.

Further example abound with more given in \cite{lewis2014}. To generalize, we write the full likelihood in two pieces
\begin{eqnarray}
\label{FullLikelihood}
L(\bth | \by)  & = & f(T(\by) | \bth) \,\, f(\by |\bth, T(\by)) .  
\end{eqnarray}
where $T(\by)$ is any statistic. Here,  $f(T(\by) | \bth)$ is the conditional pdf of $T(\by)$ given $\bth$ and $f(\by |\bth, T(\by))$ is the conditional pdf of $\by$ given $\bth$ and $T(\by)$.  In \eqref{OutlyingCases}, the conditioning statistic is $T(\by) = (y_1, \ldots, y_{n-k})$ and in \eqref{Censoring}, the conditioning statistic is $T(\by) = (c(y_1),\ldots,c(y_n))$.  We refer to 
$f(T(\by) | \bth)$ as the restricted likelihood and $L(\bth | \by)=f(\by|\bth)$ as the full likelihood.

$T(\by)$ is a well-defined random variable with a probability distribution indexed by $\bth$ and the update from prior  to posterior can be made using $f(T(\by)|\bth)$ rather than $f(\by|\bth)$.  This leads to the \textit{restricted likelihood posterior}
\begin{eqnarray}
\label{RestrictedPosterior}
\pi(\bth | T(\by)) & = & \frac{\pi(\bth) f(T(\by) | \bth)}{m(T(\by))} ,
\end{eqnarray}
where $\pi(\bth)$ is the prior distribution and $m(T(\by))$ is the marginal distribution of $T(\by)$ under prior. The associated predictive distribution for a single observation $\tilde y$ is
\begin{eqnarray}
\label{RestrictedpredDist}
f(\tilde y | T(\by)) = \int f(\tilde y | \bth) \pi(\bth | T(\by))\ d\bth .  
\end{eqnarray}

Direct use of restricted likelihood appears in many areas of the literature.  The motivation is often similar to ours:   
concern about outliers or, more generally, model misspecification.  The use of rank likelihoods is discussed by \cite{savage1969}, \cite{pettitt1983, pettitt1982}, and more recently by \cite{hoff2013}.  \cite{lewis2012} make use of robust regression estimators for $T(\by)$.
Asymptotic properties of restricted posteriors are studied by \cite{doksum1990}, \cite{clarke1995}, \cite{yuan2004},  and \cite{hwang2005}. These asymptotic results asserts that under certain assumptions, posterior distribution resembles the asymptotic sampling distribution of the conditioning statistic.

Restricted likelihoods have been used as practical approximations to a full likelihood. For example, \cite{pratt1965} appeals to heuristic arguments regarding approximate sufficiency to justify the use of the restricted likelihood of the sample mean and standard deviation. Approximate sufficiency is also used in Approximate Bayesian Computation (ABC), which is related to our method.  
ABC has recently experienced success in applications to epidemiology, genetics, and quality control \citep[see, for example,][]{tavare1997, pritchard1999,  marjoram2003, fearnhead2012}. Interest typically lies in the full data posterior and ABC is used for computational convenience.  Consequently, effort is made to choose an approximately sufficient $T(\by)$ and update to the ABC posterior by using the likelihood $L(\bth| \mathcal{B}(\by))$, where $\mathcal{B}(\by)=\{\by^{*}|\rho(T(\by),T(\by^{*}))<\epsilon\}$, $\rho$ is a metric, and $\epsilon$ is a tolerance level. This is the likelihood conditioned on the collection of data sets close to $\by$ as measured by the summary statistic. By choosing an approximately sufficient $T(\cdot)$ and a small enough $\epsilon$ the claim $L(\bth|\mathcal{B}(\by))\approx L(\bth|T(\by))\approx L(\bth|\by)$. Consequently, the ABC posterior approximates the full data posterior. Efforts have been made to formalize what is meant by  approximate sufficiency \citep[e.g.,][]{joyce2008}.

ABC is related to our method in that conditioning is on something other than the full data. For technical reasons, the calculation proceeding from $\pi(\bth)$ to $\pi(\bth | \mathcal{B}(\by))$ is not a formal conditional update.  Simply put, if an external observer told $\mathcal{B}(\by)$ can supply $T(\by)$ and vice-versa, the information contained in $\mathcal{B}(\by)$ is exactly that contained in $T(\by)$ and proper conditioning will lead to identical posterior distributions.  This is not the case with a typical implementation of ABC.

@@

This work extends the development of Bayesian restricted likelihood by arguing that deliberate choice of $T(\by)$ is sound practice and also by expanding the class of conditioning statistics in which exact conditioning can be achieved. Our methods do not rely on asymptotic properties, approximate conditioning, nor do they need to correspond to artificial censoring to create incomplete or missing data \citep[e.g.,][]{albert1988, hoffwakefield2013}.  \green{Note - fix references here hoffwakefield for asymptotic; albert for ? .}

>>>>

\documentclass[12pt]{article}
%\documentclass[ba]{imsart}

@@

\newcommand{\bx}{\mbox{\boldmath $x$}}
\newcommand{\bu}{\mbox{\boldmath $u$}}
\newcommand{\bc}{\mbox{\boldmath $c$}}
\newcommand{\bs}{\mbox{\boldmath $s$}}
\newcommand{\by}{\mbox{\boldmath $y$}}
\newcommand{\bz}{\mbox{\boldmath $z$}}
\newcommand{\bv}{\mbox{\boldmath $v$}}
\newcommand{\bb}{\mbox{\boldmath $b$}}
\newcommand{\bzero}{\mbox{\boldmath $0$}}
\newcommand{\thstarj}{\mbox{$\theta^\star_j$}}
\newcommand{\bths}{\mbox{$\btheta^\star$}}
\newcommand{\pkg}[1]{{\fontseries{b}\selectfont #1}}

@@

%  Points to handle in preparing the manuscript
%
%  1.  Out of sample evaluation of the model.  We should go with trimmed mean log-likelihood (trimming the cases
%       that contribute least to log-likelihood).  Justification for the trimming mentioned in a sentence
%       here, to be more carefully and completely done in separate work (with Yoonsuh).  
%
%  3.  Add quite a number of references to Section 2.  This is where we should catch the old work
%      that conditions on the sample mean and proceeds through asymptotic approximation, Peter 
%      Hoff's work on ranks, Holmes and Walker's work, and ABC.  New text needs to accompany the
%      references, making clear what the other work does and how ours differs.  The new text should
%      be quite brief.  
%
%  4.  A big question.  Should we try to add any asymptotic theory?  A Section 3.4?  See next comment.  
%      My vote is "no".  
%
%  5.  Page limitations.  A goal for the entire manuscript is 30 pages.  Current length is 34+, spilling
%      onto page 35.  But, additions will include the abstract (1/2 page or so), conclusions (1 page or 
%      so), more references (another half page?), and more writing scattered throughout.  Redoing
%      figures (two to drop) and tables will save a modest amount of space.   I propose that we 
%      hold off on adding new content until we see how long the paper is.  Also, new content along
%      the lines of point 4 would mean that we have to prove/use existing results to show the truth
%      of whatever we claim!  
%
%  6.  Confirm with Nationwide that they are okay with our using their data and that they are comfortable
%       with the descriptions we have in the paper.  
%
%  7.  Fit the hierarchical model, conditioning on the robust summaries at each leaf in the model.  
%
%  9.  A hedge on computations?  There are examples (pathological in some senses) for which a basic
%      MCMC based on the complete data works, but where conditioning on a poorly chosen T(y) leads
%      to a reducible Markov chain.  I don't think this happens when the support of y is the same for all
%      \theta (it may be that mutual absolute continuity of all distributions for y|\theta is needed, though
%      definitions of "support" vary somewhat).  I think we can ignore this here, but an example could go
%      into the dissertation.  (one sentence now included to hint at this).

%  To be added
%
%  1.  Abstract.  
%
%  2.  Conclusions.  To be written later.  I will take a pass at them once a few results are filled in for
%       the data analysis.  
%

%  John.  
%  See notes in red throughout.  The overarching goal is to move the manuscript toward completion.  
%  To do so means filling in details to get them taken care of.  Big batches are (i) references, (ii) 
%  deciding on notation for the estimator (b to replace beta?), (iii) verifying that the theorems have the
%  right conditions, (iv) patching figures, (v) adding material for the examples, and (vi) deciding whether
%  the figure and text on the last two pages is needed.  
%  If you want to add comments for us, pick a color other than red and add them in that color.

%  Yoon.  
%  A check on the theorems would be great.  No need, I think, to dig out the paper by Miao and Ben Israel
%  as that part of the math is fine (I'll check to make sure that all conditions for our use of the theorems have
%  been stated in the Corollary).  Once more details are in place, proof reading and adjustment.

%\newpage
\title{Bayesian Restricted Likelihood Methods: Conditioning on Insufficient Statistics in Bayesian Regression}
\author{John R. Lewis, Steven N. MacEachern and  Yoonkyung Lee \\
{\small \it Department of Statistics, The Ohio State University, Columbus, Ohio 43210}\\
{\small lewis.865@osu.edu, snm@stat.osu.edu and yklee@stat.osu.edu}
\thanks{This research has been supported by Nationwide Insurance Company and by the NSF under grant numbers DMS-1007682 and DMS-1209194.  The views in this paper are not necessarily those of Nationwide Insurance or the NSF.}} 
\begin{document}
\date{}
\maketitle

\begin{abstract}
Bayesian methods have proven themselves to be successful across a wide
range of scientific problems and have many well-documented advantages
over competing methods. However, these methods run into difficulties
for two major and prevalent classes of problems: handling data sets
with outliers and dealing with model misspecification. We outline the
drawbacks of previous solutions to both of these problems (e.g., use of
heavy-tailed likelihoods) and propose a new method as an
alternative.  When working with the new method, we summarize
the data through a set of insufficient statistics,
targeting inferential quantities of interest, and update the prior
distribution with the summary statistics rather than the complete
data.  By careful choice of conditioning statistics, we
retain the main benefits of Bayesian methods while reducing the
sensitivity of the analysis to features of the data not captured by
the conditioning statistics. For reducing sensitivity to outliers,
classical robust estimators (e.g., M-estimators) are natural choices
for conditioning statistics. With these choicees, the method can be thought of as a
blend of classical robust estimation and Bayesian methods.

A major contribution of this work is the development of a data 
augmented Markov chain Monte Carlo (MCMC) algorithm
for the linear model and a wide range of choices for
summary statistics. We demonstrate the method on an
insurance agency data set containing many outliers and subject to model
misspecification. Success is manifested in better predictive
performance for data points of interest as compared to competing
methods.

@@

\section{Introduction}
Bayesian methods have provided successful solutions to a wide range of scientific problems, with their value
having been demonstrated both empirically and theoretically.  
In simple settings, the success of the methods is often attributed to formal optimality properties, sometimes derived
through the laws of subjective probability 
and sometimes through admissibility and the complete class
theorems of decision theory.  In complex settings, the hierarchical model allows one to create 
and fit sophisticated models that may, for example, pool information across similar problems.

The development of Bayesian inference relies on a complete Bayesian model consisting of 
three elements:  the prior distribution, the loss function, and the likelihood or sampling density.  While formal
optimality of Bayesian methods is unquestioned if one accepts the validity of all three of these elements, 
 a healthy skepticism encourages us to question each of them.  Concern about the prior distribution has
been addressed through the development of techniques for subjective elicitation \citep{garthwaite2005, ohagan2006} \blue{(added the O'Hagan text here)} and
the rise of objective Bayesian methods \citep{berger2006}.  
Concern about the loss function is reflected in, for example, the
extensive literature on Bayesian hypothesis tests \citep{kass1995}.  
The sampling density has been given less attention from a specifically
Bayesian view, although the work on predictive diagnostics \citep{box1980} departs from classical
traditions.

%The focus of this work is the intersection of Bayesian methodology and data analysis. In particular, we develop techniques to handle imperfections in the sampling density
The focus of this work is the development of techniques to handle imperfections in the sampling density.  These imperfections often show
themselves through the presence of outliers--cases %--here taken to be 
not reflecting the phenomenon under
study. %--in the data set. 
%A traditional solution to Bayesian outlier-handling involves replacing the original sampling density with a mixture or heavy-tailed density.  A mixture model  includes one component for the ``good'' data and a second component for the ``bad'' data in an attempt to compartmentalize the outliers.  The heavy-tailed approach simply attempts to discount the outliers directly. Another traditional approach fits a flexible (typically nonparametric) model to the data, producing a Bayesian version of a density estimate for both good and bad data.  In recent development, inference is made through the use of robust inference functions \citep[][in press]{lee2013}.  
%\red{Q: Is the above sentence still about modifying the sampling density to
%  handle outliers?}
There are three main solutions to Bayesian outlier-handling.  The first is to replace the basic sampling
density with a mixture model which includes one component for the ``good'' data and a second 
component for the ``bad'' data.  With this approach, the prior distribution is updated with the likelihood
from the mixture model to obtain the complete-data posterior distribution.  The good component of
the sampling density is used for prediction of future good data.  The second approach replaces the
basic sampling density with a thick-tailed density in an attempt to discount outliers, yielding techniques that 
often provide solid estimates of the center of the distribution but 
do not easily translate to predictive
densities for further good data.  The third approach fits a flexible (typically nonparametric) model to 
the data, producing a Bayesian version of a density estimate for both good and bad data.  In recent 
development, inference 
is made through the use of robust inference functions \citep[][in press]{lee2013}.

The traditional strategies for handling outliers all have their drawbacks.  While we view the sampling 
density for the good data as stable, the outlier-generating processes 
may be transitory in nature, constantly shifting as the source of bad data changes.  This prevents
us from appealing to large-sample arguments to claim that, with enough data, we can nail down a
model for both good and bad data combined.  Instead of attempting to model both good and bad
data, we propose a novel strategy for handling outliers. %:  the use of restricted likelihood.  
In a nutshell, we begin with a complete model   
as if all of the data are good.
But rather than driving the move from prior distribution to posterior
distribution by the entire likelihood, we use only the likelihood
driven by a
few summary statistics which typically target inferential quantities
of interest.  We call this reduced likelihood a restricted % an incomplete
likelihood. 
%\red{[better to call a restricted likelihood? if we change, make sure to change consistently throughout the paper]} 
%\green{[I agree with Yoon.  Better to call this a restricted likelihood.]}
The 
update is a formal update from prior distribution to posterior distribution, based on the sampling density
of the summary statistics. 
 %In our approach, the reader will identify a stream of reasoning which is manifested in classical M-estimation, generalized estimating equations, approximate Bayesian computation, and elsewhere.  
The novelty of the work is twofold.  We make use of
classical robust estimators as  summary statistics in a formal
Bayesian framework, using the sampling density of the estimators as a replacement for the 
sampling density of the data.  Second, we advance the argument that conditioning on an insufficient
summary of the data is sound practice, rather than merely being done for computational and 
modelling convenience.

% \red{[Steve, just a reminder that a paragraph about model misspecification was here orginally but commented out by John following a referee's suggestion. I also agree with the referee in that model misspecification could be a much broader issue than having a fraction of bad data while the model for good data is correctly specified.]}
%In addition to outlier-prone data, the 
\green{ The sampling density can err due to other forms of model misspecification.  
The traditional view is that, if the model is inadequate, one should build a better model.  In our
empirical work, as data sets have become larger and more complex, we have bumped into 
settings where we cannot realistically build the perfect model.  We ask the question ``by attempting
to improve our model through elaboration, will the overall performance of the model suffer?''  If yes,
we avoid the elaboration, retaining a model with some level of misspecification.  Acknowledging that
the model is misspecified implies acknowledging that the sampling density is incorrect, as
we do when outliers are present.  Thus, outliers are but one form of misspecified model.  
We recommend a common solution for dealing with outliers and more general forms of model misspecification.}

The remainder of the paper develops Bayesian restricted likelihood (Section~\ref{restrictedlikelihood}), shows
how it can be applied to a Bayesian linear model (Section~\ref{BayesLinMod}), illustrates its use
on an insurance agencies data set with a novel twist on model evaluation (Section~\ref{Applications}), 
and wraps up with a discussion (Section~\ref{Conclusions}).  
A major contribution of this work is the computational strategy whose legitimacy is established 
in Section~\ref{BayesLinMod}.  The technical proofs are in the appendix.

\section{Restricted Likelihood}
\label{restrictedlikelihood}
%\textcolor{blue}{Tried to make more concise. Commented out section describing the further examples related to randomized experiments, contingency tables, and meta-analysis. }

\subsection{Examples}
To describe the use of the restricted likelihood in a Bayesian framework, 
we begin with a pair of simple examples
for the one-sample problem. Several more examples are detailed in \cite{lewis2014}.  For those discussed here, the model takes the
data $\by=(y_1,\ldots,y_n)$
to be a random sample
of size $n$ from a continuous distribution indexed by a parameter
vector $\bth$, with pdf $f(y| \bth)$.  The standard, or full-data,
likelihood would be $L(\bth | \by) = \prod_{i=1}^n f(y_i | \bth)$.

As a first example, we consider the case where a subset of the data are known to be 
bad in the sense of not informing us about $\bth$--and where the subset is known.  This case mimics the setting where outliers in
a data set are identified and discarded before a formal analysis is done.  
Without loss of generality, we label the good 
cases $1$ through $n-k$ and the bad 
cases $n-k+1$ through $n$.  
The relevant likelihood to be used to move from prior distribution to posterior distribution is clearly 
$L(\bth | y_1, \ldots, y_{n-k}) = \prod_{i=1}^{n-k} f(y_i | \bth)$.  
For an equivalent analysis, we rewrite the full %entire
likelihood as the product of two pieces,
\begin{eqnarray}
\label{OutlyingCases}
L(\bth | \by)  
= \left( \prod_{i=1}^{n-k} f(y_i | \bth) \right) \left( \prod_{i=n-k+1}^{n} f(y_i | \bth) \right) .  
\end{eqnarray}
% where the second factor may not actually involve $\bth$.
We wish to keep the first piece and drop the second for better inference on $\bth$.

A second example involves deliberate censoring of small and large observations.  This is
sometimes done as a precursor to the analysis of reaction time
experiments  \citep[e.g.,][]{ratcliff1993}
\green{where very small reaction times are physiologically too quick to be true reaction times
are usually explained as anticipation, and where implausibly long reaction times are believed due to inattention. }  
%\red{Q: provide a reason why this deliberate censoring is done?}
With lower
and upper censoring times at $t_1$ and $t_2$, the post-censoring sampling distribution
is of mixed form, with masses $F(t_1|\bth)$ at $t_1$ and $1-F(t_2|\bth)$ at $t_2$,
and density $f(y | \bth)$ for $y \in (t_1, t_2)$.  We adjust the original data $y_i$,
producing $c(y_i)$ by defining $c(y_i)= t_1$ if $y_i \leq t_1$, $c(y_i)=t_2$ 
if $y_i \geq t_2$, and $c(y_i)=y_i$ otherwise.  
The adjusted update is performed with $L(\bth |c(\by))$.  
With slightly non-standard notation, we let $g(t_1|\bth) = F(t_1|\bth)$,
$g(t_2 | \bth) = 1 - F(t_2|\bth)$, and $g(y|\bth)=f(y|\bth)$ for
$y \in (t_1, t_2)$.
Hence, we may rewrite the full %complete %entire 
likelihood as the product of
two pieces, 
\begin{eqnarray}
\label{Censoring}
%L(\bth | \by) =  \left( \prod_{i=1}^n g(x_i | \bth) \right) \left( \prod_{i=1}^n f(y_i|\bth, x_i) \right) ,
L(\bth | \by) =  
\left( \prod_{i=1}^n g(c(y_i)  | \bth) \right) \left( \prod_{i=1}^n f(y_i|\bth,c(y_i) ) \right) ,  
\end{eqnarray}
and retain only the first for the formal update.

%Further examples abound. In a completely randomized experimental design, we randomize %our 
%experimental units to treatment conditions and then ignore the details of the observed randomization \citep{dean1999}; 
%in work with contingency tables,  
%we collapse categories with small counts, coarsening the scale of %our 
%data \citep{agresti2002}; in meta-analysis, we ignore the individual patient-level data and instead
%work with estimated effects from the studies \citep{orourke2007}.  
%Further examples are described in \cite{lewis2014}. 
\subsection{Generalization}

To generalize the approach in \eqref{OutlyingCases} and
\eqref{Censoring}, %, and these other settings,
we write the full likelihood in two pieces with a conditioning statistic $T(\by)$, as indicated below:
\begin{eqnarray}
\label{FullLikelihood}
L(\bth | \by)  
& = & f(T(\by) | \bth) \,\, f(\by |\bth, T(\by)) .  
\end{eqnarray}
%Throughout the paper, we use the convention that a function is defined by its inputs. 
Here,  $f(T(\by) | \bth)$ is the conditional pdf of $T(\by)$ given $\bth$ and $f(\by |\bth, T(\by))$ is the conditional pdf of $\by$ given $\bth$ and $T(\by)$.  In the dropped case example, the conditioning statistic is $T(\by) = (y_1, \ldots, y_{n-k})$.  In 
the censoring example, the conditioning statistic is $T(\by) = (c(y_1),\ldots,c(y_n))$.  We refer to 
$f(T(\by) | \bth)$ as the restricted likelihood and $L(\bth | \by)=f(\by|\bth)$ as the full likelihood.

Bayesian methods can make use of a restricted likelihood %in place of a complete likelihood
since $T(\by)$ is a well-defined random variable with a probability distribution indexed by $\bth$.  
The update from prior distribution to posterior distribution can be 
made on the basis of $f(T(\by)|\bth)$ rather than $f(\by|\bth)$.  This leads to the restricted likelihood posterior 
\begin{eqnarray}
\label{RestrictedPosterior}
\pi(\bth | T(\by)) & = & \frac{\pi(\bth) f(T(\by) | \bth)}{m(T(\by))} ,
\end{eqnarray}
where $m(T(\by))$ is the marginal distribution of $T(\by)$ under the prior distribution.  
Predictive statements for further (good) data rely on the model.  For another observation,
say $y_{n+1}$, we would have the predictive density 
\begin{eqnarray}
\label{RestrictedpredDist}
f(y_{n+1} | T(\by)) = \int f(y_{n+1} | \bth) \pi(\bth | T(\by))\ d\bth .  
\end{eqnarray}

\subsection{Literature review}
Direct use of restricted likelihood appears in many areas of the literature.  The motivation is often similar to ours:   
concern about outliers or, more generally, model misspecification.  For example, the use of rank likelihoods is discussed by \cite{savage1969}, \cite{pettitt1983, pettitt1982}, and more recently by \cite{hoff2013}.  
\green{Lewis et al. (2012 xxxx) make use of robust regression estimators for $T(\by)$.}  
Asymptotic properties of restricted posteriors are studied by \cite{doksum1990}, \cite{clarke1995}, \cite{yuan2004},  and \cite{hwang2005}. The tenor of these asymptotic 
results is that, for a variety of conditioning statistics with non-trivial regularity conditions on prior, model, and likelihood, the
posterior distribution resembles the asymptotic sampling distribution of the conditioning statistic.

%edited to include more detail on ABC:
%\textcolor{blue}{Part on the use of incomplete likelihoods being used as approximations to complete likelihoods was commented out. However, I added more detail on ABC. Thought is that we could move this to the discussion.} 
%\red{This additional discussion on ABC
%  (next two paragraphs) looks a bit too long.}
Restricted likelihoods have also been used as practical approximations to a full likelihood. For example, \cite{pratt1965} appeals to heuristic arguments regarding approximate sufficiency to justify the use of the restricted likelihood of the sample mean and standard deviation. Approximate sufficiency is also appealed to in the use of Approximate Bayesian Computation (ABC), which is related to our method.  
ABC is a collection of posterior approximation methods which has recently experienced success in applications to epidemiology, genetics, and quality control \citep[see, for example,][]{tavare1997, pritchard1999,  marjoram2003, fearnhead2012}. Interest typically lies in the full data posterior and ABC is used for computational convenience as an approximation.  Consequently, effort is made to choose an approximately sufficient $T(\by)$ and update to the ABC posterior by using the likelihood $L(\bth| \mathcal{B}(\by))$, where $\mathcal{B}(\by)=\{\by^{*}|\rho(T(\by),T(\by^{*}))<\epsilon\}$, $\rho$ is a metric, and $\epsilon$ is a tolerance level. This is the likelihood ``conditioned'' on the collection of data sets which result in a $T(\cdot)$ within $\epsilon$ of the observed $T(\by)$. %Hence, we can see that, as in our method, conditioning is done on a summary of the complete data. 
By choosing an approximately sufficient $T(\cdot)$ and a small enough $\epsilon$ the claim is made that $L(\bth|\mathcal{B}(\by))\approx L(\bth|T(\by))\approx L(\bth|\by)$. Consequently, the ABC posterior approximates the full data posterior. We note that efforts have been made to formalize what is meant by  approximate sufficiency \citep[e.g.,][]{joyce2008}.

%ABC is related to our method in that conditioning is on something other than the complete data. Even if $T(\by)$ were exactly  sufficient in the traditional sense, a positive $\epsilon$ would result in something theoretically different from the complete data posterior (even if they are indeed approximately the same). We also mention that in ABC, $\mathcal{B}(\by)$ is not a statistic in the strict sense because it does not partition the sample space before observing data. That is, a feasible data set $\bz^{*}$ can belong to more than one `ball'  $\mathcal{B}(\cdot)$.  The method relies on conditioning on this set because in most ABC applications, approximating the complete data posterior directly is hard (or impossible) and sampling data matching the observed summary exactly is also hard (or impossible). ABC methods are a practical and useful solution in these settings. For the regression models we present, sampling new data matching the observed summary exactly is indeed possible and so we can condition exactly on the observed summary. Further, we reiterate that our motivation is different from ABC in that we do not want to condition on the complete data because we assume it contains inherent flaws which can be detrimental to inference.

\green{ABC is related to our method in that the ``conditioning'' is on something other than the data $\by$.  However, the conditioning in ABC is heuristic.  For technical reasons, the calculation proceeding from $\pi(\bth)$ to $\pi(\bth)L(\bth| \mathcal{B}(\by))/\int \pi(\bth)L(\bth| \mathcal{B}(\by)) d\bth$ is not a formal conditional update.  Simply put, if an external observer told $\mathcal{B}(\by)$ can supply $T(\by)$ and vice-versa, the information contained in $\mathcal{B}(\by)$ is exactly that contained in $T(\by)$ and proper conditioning will lead to identical posterior distributions.  This is not the case with a typical implementation of ABC.}

@@

This work extends the development of Bayesian restricted likelihood by arguing \green{that deliberate choice of $T(\by)$} is sound practice and also by expanding the class of conditioning statistics in which exact conditioning can be achieved. %well beyond ranks.  
Our methods do not rely on asymptotic properties, \green{nor do they rely on} approximate conditioning, \green{nor do they need to correspond to artificial censoring to create incomplete or missing data}  \citep[e.g.,][]{albert1988, hoffwakefield2013}.  
\green{Note - fix references here hoffwakefield for asymptotic; albert for ? .}
